---
title: Make your own QR code generator with Shiny
author: An Chu
date: '2018-05-08'
slug: make-your-own-qr-code-generator-app-with-shiny
categories: []
tags:
  - shiny
  - R
draft: true
---

QR code is a great design, especially when we consider its initial
[purpose](http://www.qrcode.com/en/history/) which was designed as a replacement
of linear barcodes - a storage medium of commodity prices in supermarkets. These
days, the two-dimensional barcode has been widely
[adopted](https://a16z.com/2017/08/11/qr-codes-in-china/) in many other
industries, notably manufacturing, marketing, online payment, and that goes well
beyond its creator's intention.

At [RTA](https://cms.rta.vn/), we utilize QR code at many places, from verifying
employee's daily check-in check-out to storing and retrieving company's
assets. Equally important, it makes our mobile-based solution to food tracking
system feasible. Given that needs, we all agree to build a comprehensive service
to generate, distribute and archive QR codes. In what follows, I will talk about
the first part which is creating a web-based QR code generator with the
excellent [Shiny](https://shiny.rstudio.com/) package.

### From string to QR code

Let's start with a simple use case: turning a string into a QR code. In the R
world, you can (safely) assume that there might be someone else already did
that. Well, it's true. Bob Rudis, a prolific R blogger, has done a terrific job
with package [`qrencoder`](https://github.com/hrbrmstr/qrencoder), bridging R
interface to `libqrencode` - a high performance C library for matrix barcode
generation. To demonstrate, I copy a toy example as follows:

```{r, eval=FALSE}
## devtools::install_github("hrbrmstr/qrencoder")
library(qrencoder)
par(mar=c(0,0,0,0))
image(qrencode_raster("QRCODE"),
      asp=1, col=c("white", "black"), axes=FALSE,
      xlab="", ylab="")
```

Basically, two things happened. First, the C library carried out the heavy work
of encoding the string to a matrix of zeroes and ones (the encoding method is
out of scope of this post, see
[here](https://www.thonky.com/qr-code-tutorial/data-encoding) for details). Then
R in turn ploted the matrix as an image of (seemingly) disordered black square
boxes. Thanks to the high level implementation of `image()` (a S4 method of
class `Raster Layer`), we are kept way from the nitty-gritty of matrix
manipulation.

Making only one or a few QR codes in R as above might be an overkill because
there're hundreds of
[online](https://duckduckgo.com/?q=online+qrcode+generator+&t=lm&ia=web) tools
for doing that. Usually, we want to produce a large number of randomly unique QR
codes in order to differentiate thousands of items or processes, and it must
also be easy to layout those barcodes for printing. This is when Shiny comes to
help.

### Generate random strings

To make random strings, I'm going to prepare data and design an interface for
those inputs. The first part, the data, can be constructed from a list of
alphanumeric characters. Besides, we'll have two options to decide the length
(in characters) and the number of of strings to be generated. The second part,
the UI, for illustrative purpose, I'll embed Shiny components (widgets) in an R
Markdown file to create the UI.

```{r, eval=FALSE}
upcase_letters <- LETTERS
locase_letters <- letters
digits <- 0:9
minus <- "-"
underscore <- "_"
space <- " "
special_char <- c("!", "@", "#", "$", "%", "^", "&", "*", "?", "/", "|", ",",
                  ".", ";", ":", "+")
brackets <- c("{", "}", "[", "]", "(", ")", "<", ">")
char_set <- list(upcase_letters, locase_letters, digits, minus, underscore,
                 space, special_char, brackets)
```

For example, the following codes create two Shiny widgets.

```{r, eval=FALSE}
fluidRow(
    column(
        width = 6,
        checkboxInput(
            inputId = "upcase_ip",
            label = "Upper-case (A, B, C,...)",
            value = TRUE
        ),
        checkboxInput(
            inputId = "locase_ip",
            label = "Lower-case (a, b, c,...)",
            value = TRUE
        ),
        checkboxInput(
            inputId = "digit_ip",
            label = "Digits (0, 1, 2,...)",
            value = TRUE
        ),
        checkboxInput(
            inputId = "minus_ip",
            label = "Minus (-)",
            value = FALSE
        )
    ),
    column(
        width = 6,
        checkboxInput(
            inputId = "underline_ip",
            label = "Underline (_)",
            value = FALSE
        ),
        checkboxInput(
            inputId = "space_ip",
            label = "Space ( )",
            value = FALSE
        ),
        checkboxInput(
            inputId = "special_ip",
            label = "Special character (!, $, %,...)",
            value = FALSE
        ),
        checkboxInput(
            inputId = "brackets_ip",
            label = "Brackets ([, ], {, },...)",
            value = FALSE
        )
    )
)
```

The UI looks like this when rendered:
